Geometric variational problems

### 2020-11-10 NSF2020/Description/background.tex
yny

\begin{wrapfigure}[16]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{The exterior domain $\Omega$ is the solvent region surrounding a
number of rigid particles $P_1,$ $P_2,$ $P_3,$ \dots 
The surface $S$ is the solvent-particle interface, indicated by the unit
normal $\mathbf{n}.$ }\label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the 
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM}.
This model addresses the major shortcomings of molecular dynamics (MD) and continuum approaches. Based on preliminary results (\S\ref{})
the PIs propose to extend this HAP model to offer
 an alternative modeling methodology that leads to new mathematical ideas,
and is both physically accurate and computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles.
These results will facilitate optimal designs of smart materials of desirable properties by
tuning the geometry and properties of the amphiphilic particles.

@@

Based on the physical origin of hydrophobicity,  we have devised a functional, called the hydrophobic attraction potential (HAP), to
model hydrophobic forces \cite{Fu2018_SIAM}. The motivation for HAP concept stemmed from PI RR's 
earlier work on the problem of energy barriers in membrane fusion \cite{RyKlYaCo16,Chetal16}.
There, PI RR and collaborators resolved the long-standing issue of accounting for the energy of a fissure surface
during topological transitions, by applying a squared gradient mathematical theory 
for hydrophobic attraction between planar surfaces \cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977} to monolayer fissures.
Based on the PI YNY's coarse-grained membrane modeling work \cite{Fu2017}, 
the investigators initiated a collaboration to create a gradient theory for arbitrary collections 
of hydrophobic and amphiphilic particles. This new method 
eliminates the costly calculation of water by treating the solvent implicitly, 
and avoids complicated re-meshing schemes of continuum approaches
by utilizing a particle-based representation.

@@

Our HAP approach differs from prior methods in a number of respects. 
First, we do not assume a surface. Rather, we more fundamentally  consider a collection of amphiphilic particles.
The collection of amphiphiles minimize hydrophobic interactions by sequestering hydrophobic tails in the form of a bilayer, and the particles' excess free energy gives rise to an elastic bilayer energy. 
The second difference lies in the fluid-interface coupling. Here, the associated mobility problem \eqref{eq:stokes} 
is more complicated than dealing with a stress boundary condition or diffusive surface force, 
because the fluid velocity must be that of a rigid body motion at the particle surfaces.

\begin{wrapfigure}[17]{l}{0.4\textwidth}
\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
\caption{\label{fig:demixing} An initial assembly of small and 
large particles spontaneous segregates into two smaller bodies. }
\end{wrapfigure}
The greatest strength of using HAP to model lipid bilayer membrane is the ability
to form discontinuities (interfacial singularity) from first physical principle without any artificial manipulation to rearrangement the interface.
%The ability to form discontinuities is perhaps the HAP method's greatest strength.
 Indeed, a strong hear flow must cause membrane rupture since the 
energy of stretching must exceed the pore nucleation and widening barriers 
for sufficiently large flow strengths. 
Some disadvantages are that the hydrophobic interaction does not constrain vesicle volume. 
Instead, changes in volume are rate limited by inter-particle spacing,
which may be undesirable in case of a mathematically strict volume constraint.
% set vesicle volume, as is done in the study of liposomes
%using auxiliary constraint equations for instance. 
%Also, the boundary integral method formulation
%relies on linearity of the Stokes equations. There has been some progress in
%boundary integral methods for the non-linear Navier-Stokes equations [ref].
%We point out that some non-Newtonian effects in polymers are a consequence of hydrophobic
%and steric molecular interactions like the ones presented in this proposal.

>>>>

\begin{wrapfigure}[16]{r}{0.22\textwidth}
\centerline{\includegraphics[width=0.22\textwidth]{figures/BG_fig1.pdf}}
\caption{The exterior domain $\Omega$ is the solvent region surrounding a
number of rigid particles $P_1,$ $P_2,$ $P_3,$ \dots 
The surface $S$ is the solvent-particle interface, indicated by the unit
normal $\mathbf{n}.$ }\label{fig:domain}
\end{wrapfigure}
Recently PIs RR and YNY developed a mathematical model called the 
hydrophobic attraction potential (HAP) \cite{Fu2018_SIAM}.
This model addresses the major shortcomings of molecular dynamics (MD) and continuum approaches. Based on preliminary results (\S\ref{sec:preliminary_work})
the PIs propose to extend this HAP model to offer
 an alternative modeling methodology that leads to new mathematical ideas,
and is both physically accurate and computationally practical.
%
%
%The word hydrophobic (water fearing) derives from the low solubility of oil (hydrocarbon solute) in water and vice versa. 
%It causes hydrophobic moieties to aggregate and cluster,
%is responsible for the adhesion between hydrophobic surfaces \cite{Ducker2016}, large contact angles on a 
%dewetting surface \cite{Arenas2019,Sandre1999}, accumulation of particles along interfaces \cite{Lee2013,Lee2014}, 
%formation of micelles and bilayers \cite{Israelachvili80}, and protein folding and membrane insertion \cite{Kabelka2018}.
%
%Lipids are amphiphilic molecules whose  structure possesses
%both hydrophobic and hydrophilic parts. 
%The amphiphilic property is what allows lipids to form the membranes and 
%compartments of living cells \cite{Israelachvili80}. 
%More specifically, a lipid consists of 
%an elongated hydrocarbon tail that is hydrophobic, attached to a polar head that is hydrophilic.
%To shield the hydrophobic tails from water, lipids self-assemble into micelles and bilayers. 
%A micelle is a spherical arrangement of lipids with tails terminating at the micelle center. 
%A bilayer consists of two layers of lipids called monolayers, where the lipid tails point 
%from the monolayer surface into the bilayer core. 
%
%The mathematical modeling of a biological membrane is a challenging problem in applied mathematics. 
%Bilayers are elastic and resist deformations like bending, twisting, and stretching.
%Their elastic deformations are well described by the theory of liquid crystals \cite{ANDRIENKO2018520}.
%Lipid bilayer membranes can also be fluidic, and the lateral translation of lipids (or any membrane bound
%proteins) couples nonlocaly to the motion of the aqueous environment \cite{MerkelSackmannEvans1989,StoneAjdari1998_JFM,OppenheimerDiamant2009_BJ,OppenheimerDiamant2011_PRL}. 
%%Finally, the membranes of cellular 
%%compartments are constantly merging and pinching as part of intracellular trafficking. 
%%Therefore monolayer surfaces undergo discontinuous deformations. 
%
%There are two prevailing approaches in membrane modeling, and each has its advantages
%and disadvantages. Molecular dynamics (MD) is to date the only tool capable of resolving granular biological details
%The computational cost of MD, however, grows with the sixth power of the sample diameter and 
%so simulations are severely limited to 
%small system sizes and short time scales \cite{DiCarlo2019}.
%The other approach, continuum mechanics, assumes smooth surfaces and can therefore model 
%realistic systems over physical times. 
%But continuum description of a biological membrane ignores the granularity of lipid molecules, and thus requires some assumptions when a
%membrane ruptures or when two membranes fuse \cite{ChKo08}. 
%%
%%mechanics presumes, rather than predicts, the nucleation of discontinuities. 
%%
%
The proposed research aims to provide fundamental understanding of the self-assembly dynamics of amphiphilic particles.
These results will facilitate optimal designs of smart materials of desirable properties by
tuning the geometry and properties of the amphiphilic particles.

@@

Based on the physical origin of hydrophobicity,  we have devised a functional, called the hydrophobic attraction potential (HAP), to
model hydrophobic forces \cite{Fu2018_SIAM}. The motivation for HAP concept stemmed from PI RR's 
earlier work on the problem of energy barriers in membrane fusion \cite{RyKlYaCo16,Chetal16}.
There, PI RR and collaborators resolved the long-standing issue of accounting for the energy of a fissure surface
during topological transitions, by applying a squared gradient mathematical theory 
for hydrophobic attraction between planar surfaces \cite{Eriksson1989,Lum1999,Menshikov2017,Marcelja1977} to monolayer fissures.
Based on PI RR and YNY's coarse-grained membrane modeling work \cite{Fu2017}, 
the investigators initiated a collaboration to create a gradient theory for arbitrary collections 
of hydrophobic and amphiphilic particles. This new method 
eliminates the costly calculation of water by treating the solvent implicitly, 
and avoids complicated re-meshing schemes of continuum approaches
by utilizing a particle-based representation.

@@

Our HAP approach differs from prior methods in a number of respects. 
First, we do not assume a surface. Rather, we more fundamentally  consider a collection of amphiphilic particles.
The collection of amphiphiles minimize hydrophobic interactions by sequestering hydrophobic tails in the form of a bilayer, and the particles' excess free energy gives rise to an elastic bilayer energy. 
The second difference lies in the fluid-interface coupling. Here, the associated mobility problem \eqref{eq:stokes} 
is more complicated than dealing with a stress boundary condition or diffusive surface force, 
because the fluid velocity must be that of a rigid body motion at the particle surfaces.
%
\begin{wrapfigure}[17]{l}{0.4\textwidth}
\centerline{\includegraphics[width=0.4\textwidth]{figures/PW_fig2.pdf}}
\caption{\label{fig:demixing} An initial assembly of small and 
large particles spontaneous segregates into two smaller bodies. }
\end{wrapfigure}
The greatest strength of using HAP to model lipid bilayer membrane is the ability
to form discontinuities (interfacial singularity) from first physical principle without any artificial manipulation to rearrangement the interface.
%The ability to form discontinuities is perhaps the HAP method's greatest strength.
For example, using the HAP model to simulate a vesicle under a shear flow that is strong enough to cause the vesicle membrane to rupture, 
we expect the HAP model to capture the reorganization of lipid molecules on the scales of membrane thickness ($\sim 5$ nm), which is a nearly impossible
task using phase-field or immerse boundary method without extreme refinement around the membrane and artificial treatment of reconnection during the topological change of an interface.

%Some disadvantages are that the hydrophobic interaction does not constrain vesicle volume. 
%Instead, changes in volume are rate limited by inter-particle spacing,
%which may be undesirable in case of a mathematically strict volume constraint.
%% set vesicle volume, as is done in the study of liposomes
%%using auxiliary constraint equations for instance. 
%%Also, the boundary integral method formulation
%%relies on linearity of the Stokes equations. There has been some progress in
%%boundary integral methods for the non-linear Navier-Stokes equations [ref].
%%We point out that some non-Newtonian effects in polymers are a consequence of hydrophobic
%%and steric molecular interactions like the ones presented in this proposal.

### 2020-11-10 NSF2020/Description/broaderimpacts.tex
yny

The proposed mathematical analysis and modeling will provide a
transformative understanding of the collective dynamics of amphiphilic
particles such as (1) their self-assembly into micelles and bilayers,
and (2) the interaction between these building blocks. Amphiphilic Janus
particles have gained increasing popularity for fabrication of smart
materials. The proposed hybrid continuum model, mathematical analysis
and numerical algorithms will have transformative impact on precision
design for specific mechanical properties of materials made of
amphiphilic nanoparticles. An important component of this proposal is
the interdisciplinary education and training of both undergraduate
students, graduate students, and postdoctoral fellows. The combination
of mathematical modeling, analysis, and scientific computing in this
project provides a compelling example of the importance of mathematics
in biophysics and engineering applications. The concepts and methods
described here go beyond the context of amphiphilic particles and lipid
molecules. They extend to other problems featuring microscopic phase
separation that leads to formation of mesoscopic domains. This situation
arises, for example, in biological development in systems biology. The
methods developed here have the potential to impact those and other
related areas in biomedicine and biotechnology.

\subsection{Educational Impacts}
\label{subsec:Educational_plans}
The proposed research will have immediate impact on undergraduate and
graduate education. At the undergraduate level, the project will involve
and support undergraduate researcher from Fordham University. The
supported summer researchers will gain valuable first-hand experience in
inherently interdisciplinary, mathematical fluid modeling and
computation. We will incorporate research into teaching, especially in
differential equation and programming courses. PI Quaife will expose
undergraduate and high school students to the reserach through the
Undergraduate Research Opportunity Program (UROP) and Young Scholar's
Program (YSP).

Over the past eight years, PI Ryham has included many undergraduates in
the execution of research initiatives, coauthoring publications. He has
a track record for supporting underrepresented groups, being called on
by the Collegiate Science and Technology Entry Program (CSTEP) to make
opportunities for promising students, for example. He has mentored two
Clare Boothe Luce Scholars, one a U.S. Marine Corp veteran, and a high
school student in the NYU GSTEM program.

The PIs will capitalize on the synergy between NJIT's doctoral program
in Mathematical Sciences and Fordham's undergraduate focus in the
Department of Mathematics. Promising, young undergraduate scientists
from the Bronx community while find a natural pipeline into graduate
studies through working directly or indirectly on this project. On the
graduate-level education, PI Young and expects to train and
support a PhD student, Yuexin Liu (who is in her third-year at NJIT) for
two more years. Yuexin Liu has past experiences in the undergraduate
training program at NJIT, and will train an undergraduate intern. The
PIs will foster vertical integration between the senior personnel, their
collaborators, postdocs, doctoral and Bachelor students, enhancing the
learning environment.
%

>>>>

The proposed mathematical analysis and modeling will provide a
transformative understanding of the collective dynamics of amphiphilic
particles such as (1) their self-assembly into micelles and bilayers, (2) the material properties of such self-assembly, 
and (3) the interaction between these building blocks. Amphiphilic Janus
particles have gained increasing popularity for fabrication of smart
materials. The proposed hybrid continuum model, mathematical analysis
and numerical algorithms will have transformative impact on precision
design for specific mechanical properties of materials made of
amphiphilic nanoparticles. An important component of this proposal is
the interdisciplinary education and training of both undergraduate
students, graduate students, and postdoctoral fellows. The combination
of mathematical modeling, analysis, and scientific computing in this
project provides a compelling example of the importance of mathematics
in biophysics and engineering applications. The concepts and methods
described here go beyond the context of amphiphilic particles and lipid
molecules. They extend to other problems featuring microscopic phase
separation that leads to formation of mesoscopic domains. This situation
arises, for example, in biological development in systems biology. The
methods developed here have the potential to impact those and other
related areas in biomedicine and biotechnology.

\subsection{Educational Impacts}
\label{subsec:Educational_plans}
The proposed research will have immediate impact on undergraduate and
graduate education. At the undergraduate level, the project will involve
and support undergraduate researcher from Fordham University. The
supported summer researchers will gain valuable first-hand experience in
inherently interdisciplinary, mathematical fluid modeling and
computation. We will incorporate research into teaching, especially in
differential equation and programming courses. PI YNY will work on simple modeling of self-assembly problems 
with both undergraduate and high school students from Newark Science Park High School. 
PI YNY has a track record of working with local high school students for their research projects before they apply for colleges.
PI BQ will expose
undergraduate and high school students to the reserach through the
Undergraduate Research Opportunity Program (UROP) and Young Scholar's
Program (YSP).

Over the past eight years, PI RR has included many undergraduates in
the execution of research initiatives, coauthoring publications. He has
a track record for supporting underrepresented groups, being called on
by the Collegiate Science and Technology Entry Program (CSTEP) to make
opportunities for promising students, for example. He has mentored two
Clare Boothe Luce Scholars, one a U.S. Marine Corp veteran, and a high
school student in the NYU GSTEM program.

The PIs will capitalize on the synergy between both FSU's doctoral program in Scientific Computing and NJIT's doctoral programs
in Mathematical Sciences and Fordham's undergraduate focus in the
Department of Mathematics. Promising, young undergraduate scientists
from the Bronx community while find a natural pipeline into graduate
studies through working directly or indirectly on this project. On the
graduate-level education, PI YNY expects to train and
support a PhD student for
two more years. 
%Yuexin Liu has past experiences in the undergraduate training program at NJIT, and will train an undergraduate intern. 
The PIs will foster vertical integration between the senior personnel, their
collaborators, postdocs, doctoral and Bachelor students, enhancing the
learning environment.
%
